### 2018-10-26 paper/data-exploration/turnover.Rmd
Add explanation & tables and use only species data (not genus)
Genus is too broad and probably not useful for the research question

---
title: "Turnover of species and interactions"
author: "Fernando Cagua"
date: "10/23/2018"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

We want to investigate whether species interact differently at the core of their range compared to the fringes of their distribution. 
One could expect species to diversify their interactions as interaction parters become less common/available. Alternatively, one could expect species to interact with only a subset of the interaction partners, as they cannot develop new interactions when environemntal conditions are close to the limits of where possitive groth rates occur.

The aim of this document is to explore whether the interaction data available in the Bascompte's Web of Life could be adequate to answer this question or not.

We start by checking wether species (and interactions) occur in enough localities to be able to detect a change across the species ranges.

## Species & interaction turnover

### Species turnover

```{r}
require(ggplot2)
drake::loadd(species_list)

plot_freq_step <- function(freq_df){
  if ("guild" %in% names(freq_df)){
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc, colour = guild)) 
  } else {
    p <- freq_df %>%
      ggplot(aes(x = n_loc_rank, y = n_loc)) 
  }
  p +
    geom_step(size = 1) +
    scale_x_log10() +
    theme_bw() +
    scale_color_brewer(palette = "Paired", 
                       labels = c("plants", "pollinators")) +
    labs(x = "rank", 
         y = "# locations") +
    theme(legend.position = c(0.99, 0.99), 
          legend.justification = c(1,1))
}

species_list %>%
  remove_unknown_species() %>% 
  get_sp_freq_by_group(sp_name) %>% 
  plot_freq_step() +
  labs(title = "number of locations per species", 
       subtitle = "turnover of species in the Web of Life database")

# species_list %>%
#   remove_unknown_species() %>% 
#   get_sp_freq_by_group(genus) %>%
#   plot_freq_step()+
#   ggtitle("number of locations per genus")
```

```{r}
species_list %>%
  remove_unknown_species() %>% 
  get_sp_freq_by_group(sp_name) %>%
  dplyr::filter(n_loc_rank <= 10) %>%
  dplyr::select(n_loc_rank, guild, sp_name, n_loc) %>%
  dplyr::arrange(guild, n_loc_rank) %>%
  knitr::kable(col.names = c("", "guild", "species", "# locations"))
```

### Interaction turnover

```{r}
require(ggplot2)
drake::loadd(interaction_list)

interaction_list %>%
  remove_unknown_species() %>% 
  get_int_freq_by_group(pla_name, pol_name) %>% 
  plot_freq_step() +
  labs(litle = "locations per interactions", 
       subtitle = "turnover of interactions in the Web of Life database")
# 
# interaction_list %>% 
#   remove_unknown_species() %>% 
#   get_int_freq_by_group(pla_genus, pol_genus) %>% 
#   plot_freq_step() +
#   ggtitle("locations per interactions (genus)")
```

```{r}
interaction_list %>%
  remove_unknown_species() %>% 
  get_int_freq_by_group(pla_name, pol_name)  %>%
  dplyr::filter(n_loc_rank <= 20) %>% 
  dplyr::select(n_loc_rank, pla_name, pol_name, n_loc) %>%
  dplyr::arrange(n_loc_rank) %>%
  knitr::kable(col.names = c("", "plant", "pollinator", "# locations"), 
               caption = 'Top 20 most "widespread" interactions')
```