### 2019-11-17 paper/draft-info.Rmd
Check spelling draft info

---
title: "Draft information"
subtitle: "version: draft-0.3"
date: "2019-11"
# bibliography: phd-bibliography/pollen-competition.bib
csl: ecology-letters.csl
output:
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
# classoption: twocolumn
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[M-]{manuscript}

---

\onehalfspacing

## Done

* Refocused the introduction
* Moved details of the taxonomic data cleaning process to supplementary materials
* Moved details of occurrence data cleaning process to supplementary materials
* Reorganised the results
* Rewrote most of the discussion
* Changed some details about how species are cleaned
* Removed species generality and the breadth of the environmental niche from the model

## To-do

* There are some warnings for when matching environmental variables to occurrences. I need to double-check this is happening correctly. 
* Refine the supplementary results. 
* Fix references