### 2020-04-18 paper/draft-info.Rmd
Update draft info

---
title: "Draft information"
subtitle: "version: draft-0.5"
date: "2020-04-18"
# bibliography: phd-bibliography/pollen-competition.bib
csl: ecology-letters.csl
output:
  bookdown::pdf_document2:
    # base_format: rticles::peerj_article
    keep_tex: yes
    number_sections: false
toc: false
geometry: 
  - textwidth=33em
  - textheight=48\baselineskip
# classoption: twocolumn
header-includes:
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{lineno}
  - \usepackage{xr}
  - \usepackage[utf8]{inputenc}
  - \newcommand{\R}[1]{\label{#1}\linelabel{#1}}
  - \newcommand{\lr}[1]{page~\pageref{#1}, line~\lineref{#1}}
  - \externaldocument[M-]{manuscript}

---

\onehalfspacing

## Done

* Included feedback to paper (thanks Jason)
* Added many references including some to GBIF datasets
* Minor adjustments to figure captions
* Tidied up the supplementary materials

## To-do

* Fix a couple of references
* Add a few references to the methods justifying occurrence cleaning approach (thanks Audrey)

## Won't-do

* Finish the manual checks for species names that we were unable to correct automatically. It takes ~5-10 mins per species and we have ~600. Given the amount of data we already have it probably won't make much of a difference.

## Comments

* Jason, one of your comments was: *"I can't remember if I've raised this before, but if species are near to the boundaries of number of partners (close to zero or close to N, where N is the number of species in a community), a random change in the number of partners is more likely to bring them towards the mean because they can't go any higher than N or lower than 0. Is there any way to deal with this?"*. That certainly might be happening for species with few partners. I try to address that in the discussion (see \lr{regression-to-the-mean}). I don't think it happens for species with many partners because very few of the species analysed are close to N (only ~2% of plants and ~3% of animals have a normalised degree > 0.5, ~0.5% of both have a normalised degree > 0.75). To avoid confusion I've removed the "regression to the mean" statements in the discussion. 
* Appart from that, I think I followed most suggestions one way or another. **Thanks!**